---
title: "R Programing Workshop"
author: 'Ebude Yolande'
output:
  pdf_document: default
  html_notebook: default
---

This worksheet is for data science and AI beginners and made for a 1:30 mins session. To begin install Rstudio \href{https://rstudio.com/products/rstudio/download/}{\textit{(download)}} or update version.
```{r}
#install.packages("installr")
#library(installr)
#updateR()
```

The following datasets would be used throughout this worksheet:

- IoT sensor dataset \href{https://drive.google.com/file/d/1xsiMGpOULuO3Ei2U4lyUpmtTLOVHaAB3/view?usp=sharing}{\textit{(download)}} 
- Grid bikeshare data \href{https://drive.google.com/file/d/1mwQV2uDxRuEwULFdVKC8H9NZXRCkARkK/view?usp=sharing}{\textit{(download)}} 

## Content
 - Dealing with columns in a dataset
 - Selecting subsets and merging of datasets.
 - Visualization
 - Functions and Reusable code

 

# 1.Dealing with columns in a dataset
Manipulating data columns, making them easy to interprete and use in analysis.  

```{r}
iot<-read.csv('iot_telemetry_data.csv')
#View(iot) #to view the entire dataset
head(iot,3)
```
Get more information on the dataset/function
```{r}
#summary(iot)

help("read.csv")
```
POSIXct function would convert the object time class to timestamp,  \href{https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/as.POSIX*}{\textit{see more}}


```{r}
iot$ts<-as.POSIXct(iot$ts,origin = '1970-01-01')
head(iot,3)
```
The decimal points on the columns with type double.

```{r}
library(help="purrr")
```


```{r}
#install.packages(purrr)
library(purrr)
iot<-modify_if(iot, ~is.numeric(.), ~round(., 4))
tail(iot,3)
```

### Exercise I: 
 
We would look at the free bike dataset.
 
 1. Import json file 
```{r}
typeof(fromJSON('free-bike-status-1.json'))
```
 

```{r}
#install.packages("jsonlite")
library("jsonlite")
BShare<-as.data.frame(fromJSON('free-bike-status-1.json'))
tail(BShare,4)

```
 
 2. Modify the header
```{r}
names(BShare)<-c('last_updated','ttl','id','name','lon','lat','reserved','disabled')
head(BShare,3)
```

 
 3. Removing unecessary info
```{r}
strtoi(str_split('bike_12749','_')[[1]][2])
```
 

```{r}
#install.packages("stringr")
library(stringr)
new<-list()
for (i in 1:dim(BShare)[1]){  #nrow(BShare)
  
  val<-strtoi(str_split(BShare$id[i],'_')[[1]][2])
  
  new<-append(new,val)
  
}
BShare$new_id<-unlist(new)
head(BShare,3)
```

4. Save in excel 
```{r}
#install.packages("writexl")
library("writexl")
BShare$rec_update<-iot[1:21,'ts']
write_xlsx(BShare, 'bike-share.xlsx')

write_xlsx(iot[1:100,], 'iot.xlsx')
```
 
### Challenge I:

1. Import bike-share excel file, then put the last_updated column to a readable datetime value. 
2. Using the iot excel dataset, modify the device column by removing the colon that separate each term (for example 1c:bf:ce:15:ec:4d becomes 1cbfce15ec4d).

# 2.	Selecting subsets and merging of datasets.
Selecting the appropraite subset of a data to use as well as merging different datasets are very important.

```{r}
iot[5:10, 3:ncol(iot)]
```


From the iot_telemery csv dataset, we would select the data of device b8:27:eb:bf:9d:51.
```{r}
iot_d1<-iot[which(iot$device=='b8:27:eb:bf:9d:51'),]
head(iot_d1,3)
```
We can add more conditions
```{r}
iot_d1<-iot[which(iot$device=='b8:27:eb:bf:9d:51' & iot$co >0.005), c('ts','device','co','temp')]
head(iot_d1,3)
```

If just device b8:27:eb:bf:9d:51 is known and we dont want its information in the dataset

```{r}
#install.packages("dplyr")
library(dplyr)
iot_d2<-filter(iot[ , c('ts','device','co','temp')],iot$device!='b8:27:eb:bf:9d:51')
tail(iot_d2,3)
```

We would merge the two new datasets to reconstruct the old dataset.

```{r}
iot_d<-rbind(iot_d1,iot_d2)
iot_d<-bind_rows(iot_d1,iot_d2)
head(iot_d,3)
```

### Exercise II:

1. Creating a vector from taking the last 30 elements in a column ts of iot_d2 and assign to first 30 elements in iot_d1, for example
\begin{equation*}
x=[2,6,4,8,1,5,8,9],\\
new_x=[9,8,5,1]
\end{equation*}
```{r}
iot_d1[1:30,'ts']<-rev(tail(iot_d2$ts,30))
head(iot_d1,3)
```

2.Create a new dataset of iot data merging iot_d1 and iot_d2 based on column ts.
```{r}
iot_d<-merge(iot_d1,iot_d2, by='ts')
head(iot_d,3)
```

3. Select the iot data which was recorded between 12/07/2020 to 15/07/2020.
```{r}
iot_3d<-iot[which(iot$ts > '2020-07-12 00:00:00' & iot$ts < '2020-07-16 00:00:00'),]
head(iot_3d,3)
```
 

### Challenge II

1. Remove the columns last_updated, id, and ttl from the bike share excel dataset.
2. Merge the new bike-share.xlsx and iot.xlsx depending on time
3. Create a new dataset which is the subset of the one in 2. where the co value is greater than the mean value of the iot.xlsx dataset.

# 3. Visualization
We will start to visualize iot_telemetry, we start with one variable.

```{r}
hist(iot$co)
```


```{r}
#install.packages("ggplot2")
library(ggplot2) 
qplot(ts,co,data=iot,geom='line',color=device,main = "CO released by sensors")
```


### Exercise III

1. Visualize the relationship between humidity and temperature 
```{r}
png('image.png')
qplot(temp, humidity, data=iot,fill=device, color=device,main = "Humidity at various temperatures")

```


2. Distribution/Correlation  plots
```{r}
par(mfrow=c(1,2))

dist_p<-ggplot(iot,aes(device,smoke))+geom_boxplot() + theme_bw() + xlab("Device")+ ylab("Smoke") + 
  ggtitle('Boxplot of smoke emission of devices')


corr_p<- ggplot(iot, aes(smoke, co))+geom_line()+
  theme(panel.grid.minor = element_blank(), panel.background = element_blank(), 
        axis.line = element_line(colour = "black")) + xlab("Smoke")+ ylab("CO2 amount") +
  ggtitle('Correlation of smoke and Co2 emission')

ggsave('image2.png')
```


### Challenge III:

1. Using iot_telemetry dataset, visualize the distribution of smoke of each device.
2. Visualize the co emission of the devices at different temperature for the days of the week.


# 4. Functions and Reusable code

Function

\begin{equation*}
x=\frac{-b \pm \sqrt(b^2 -4ac)}{2a}
\end{equation*}

```{r}
Quad <- function(a,b,c){
  if (b^2-4*a*c >=0){
    return(c((-b-sqrt(b^2-4*a*c))/(2*a),(-b+sqrt(b^2-4*a*c))/(2*a)))
  }
  else{
    A=-b/(2*a)
    B= sqrt(-(b^2-4*a*c))/ (2*a)
    return(c('complex', A, B))
  }
}
bet<-Quad(2,1,0)
bet
```

Process
```{r}
con_json_xlsx<- function(js){
  library(jsonlite)
  library("writexl")
  dtf<-as.data.frame(fromJSON(js))
  write_xlsx(dtf,'new_name.xlsx')
}
con_json_xlsx('free-bike-status-1.json')
```

Importing another R code and calling a function from it
```{r}
source('doc.R')
Quad2(2,2,0)
```



